submission

### 2024-03-09 submission.Rmd

### 2.1 We first want to study the gene which is the most correlated with the level of cholesterol. Identify this gene and plot `cholesterol` as a function of this gene expression level. Fit a linear model. Test if there is a significant relation between this gene expression and `cholesterol`?

@@

### 2.2 We now wish to perform the previous test for all of the 3116 genes. For each of the genes, fit a linear model that explains `cholesterol` as a function of the gene expression and compute the p-value of the test. AND

### 2.3 Order the p-values and plot the ordered p-values as a function of their rank. On the same plot, display the line $y = x/3116$. Discuss the result.

@@

@@

@@

@@

##2.5 A stronger guarantee would be that on impose that there shouldn't be more than a 5 % risk that any discovery is false. For this method, we use the Holm–Bonferroni method. The method is as follows:\\

I have my ordered p-values, $\{ P_1, P_2, ... \: , P_m \}$, and want the FWER to be no higher than a level $\alpha$. Is $P_1 < \alpha/m$? so, reject $H_1$ and continue to the next step, otherwise EXIT. Is $P_2 < \alpha/(m-1)$? so, reject $H_2$ and continue to the next step, otherwise EXIT.

Continue this iterative process as long as $P_k < \frac{\alpha}{m+1-k}$. Using this process, we get the following result:

>>>>

## 2.1 Finding most significant discovery

**Task description:**: We first want to study the gene which is the most correlated with the level of cholesterol. Identify this gene and plot cholesterol as a function of this gene expression level. Fit a linear model. Test if there is a significant relation between this gene expression and cholesterol?

We first want to study the gene which is the most correlated with the level of cholesterol. Identify this gene and plot `cholesterol` as a function of this gene expression level. Fit a linear model. Test if there is a significant relation between this gene expression and `cholesterol`?

@@

## 2.2 Calculating all p-values

**Task description:**: We now wish to perform the previous test for all of the 3116 genes. For each of the genes, fit a linear model that explains cholesterol as a function of the gene expression and compute the p-value of the test.

We now do the same test for all 3116 genes and fit a linear model that explains `cholesterol` as a function of the gene expression. Doing so gives us a relatively large proportion of genes that correspond to a p-value of less than 0.05

@@

```{r}
plot(hist(p_list, breaks=20, col = c("orange", rep("lightblue", 100))), main = "Density of p-values", xlab = "p-values")

## 2.3 Ordering p-values

**Task description:**Order the p-values and plot the ordered as a function of their rank. On the same plot, display the line y=x/3116. Discuss the result.

Plotting all values, we can conclude that there are amount of models that have a p-value below 0.05. It is tempting to make the conclusion that all these models are statistically significant since they all have a small p-value. However, in the event that the we only have random data, we still expect 5% of the models to appear significantly significant. In fact, if we were to only have random data, we expect a uniform distribution between 0 and 1 for the p-values. Visually, looking at the figure above, we can the p-values perform better that the uniform distribution and therefore appear to contain significant discoveries.

@@

plot(hist(p_list_shuffeled, breaks=20, col = c("orange", rep("lightblue", 100))), main = "Density of p-values", xlab = "p-values")

@@

## 2.4 Benjamini-Hochberg procedure

**Task description:**Identify a set of genes linked to the response (aka discoveries). We want to guarantee that the expected proportion of false discoveries (mistakes) is less than 5%. Explain how you proceed and how many genes you discover.

@@

##2.5 Holm–Bonferroni method **Task description:** We wish to be more conservative and guarantee that the probability of making a false discovery (or more) is less than 5%. Explain how you proceed and how many genes you discover.

A stronger guarantee would be that on impose that there shouldn't be more than a 5 % risk that any discovery is false. For this method, we use the Holm–Bonferroni method. The method is as follows:

I have my ordered p-values, $\{ P_1, P_2, ... \: , P_m \}$, and want the FWER to be no higher than a level $\alpha$.

1.  Is $P_1 < \alpha/m$? Reject $H_1$ and continue to the next step, otherwise EXIT.

2.  Is $P_2 < \alpha/(m-1)$? so, reject $H_2$ and continue to the next step, otherwise EXIT.

3.  Continue this iterative process as long as $P_k < \frac{\alpha}{m+1-k}$.

    Using this process, we get the following result:
